---
title: "HateCrime"
author: "Roberto Conti"
date: "02/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# USE DURATION OR PERIOD forx exampale avg number of day between two crimes per eache reagion, show as table
# wider
# create join after filtering 'find regione wehich never had crimes certain type

```{r}
# Importing Library

library(tidyverse)
library(readxl)
library(lubridate)
```


# Import  and cleaning the dataset
During my cleaning:
    I have dropped column that:
       - I do not need in my analyis
       - Column with just NA values
       - Duplicate columns
       
```{r}

#overview about the columns present in the imported dataSet
 prova <- read_csv("hate_crime.csv") %>% str()

prova
#Just Select the column that I need 
 hate_crimes <- read_csv("hate_crime.csv", 
         col_select = c(
                        "STATE_ABBR", 
                        "STATE_NAME","DIVISION_NAME",
                        "REGION_NAME",
                        "POPULATION_GROUP_DESC",
                        "INCIDENT_DATE",
                        "TOTAL_OFFENDER_COUNT",
                        "OFFENDER_RACE",
                        "VICTIM_COUNT",
                        "OFFENSE_NAME",
                        "LOCATION_NAME",
                        "BIAS_DESC",
                        "VICTIM_TYPES",
                        "MULTIPLE_OFFENSE",
                        "MULTIPLE_BIAS")) %>%   # formatting data types
mutate(INCIDENT_DATE = as_date(dmy(INCIDENT_DATE))) %>% #formatting factors
   mutate(STATE_ABBR = as_factor(STATE_ABBR),
          STATE_NAME = as_factor(STATE_NAME),
          DIVISION_NAME = as_factor(DIVISION_NAME),
          REGION_NAME = as_factor(REGION_NAME),
          POPULATION_GROUP_DESC = as_factor(POPULATION_GROUP_DESC),
          OFFENDER_RACE = as_factor(OFFENDER_RACE),
          OFFENSE_NAME = as_factor(OFFENSE_NAME),
          LOCATION_NAME = as_factor(LOCATION_NAME),
          BIAS_DESC = as_factor(BIAS_DESC),
          VICTIM_TYPES = as_factor(VICTIM_TYPES),
          MULTIPLE_OFFENSE = as_factor(MULTIPLE_OFFENSE),
          MULTIPLE_BIAS = as_factor(MULTIPLE_BIAS),
          ) %>% drop_na(OFFENDER_RACE)
  
 

```

 Total Number of Hate Crime Cases and  each Year
```{r}
hate_crimes %>%
  group_by(year = year(INCIDENT_DATE), REGION_NAME) %>% 
  summarise(totalHatecrimes = n()) %>% 
  ggplot(aes(year, totalHatecrimes, color = REGION_NAME)) + geom_line()


hate_crimes %>%
  group_by(year = year(INCIDENT_DATE)) %>% 
  summarise(totalHatecrimes = n()) %>% 
  ggplot(aes(year, totalHatecrimes)) + geom_line()

```
Show total number of hate crimes, offenders and victims for each Region

```{r}
# I want plot all of thre on y axis. how?

ggplotly(hate_crimes %>% 
  group_by(REGION_NAME) %>% 
  summarise(TotHateCrimes = n(),
            TotOffenders = sum(TOTAL_OFFENDER_COUNT),
            TotVictims = sum(VICTIM_COUNT)) %>% 
  pivot_longer(cols = c('TotHateCrimes', 'TotOffenders', 'TotVictims'),
               names_to = "Type", values_to = "total") %>% 
  ggplot(aes(fct_reorder(REGION_NAME, total,.desc = TRUE), total, fill = as.factor(Type))) + geom_col(position = "dodge") )

```




Collapse levels of 'BIAS_DESC'
```{r}


#remove multiple BIAS from 'BIAS_DESC' column, because I am interest in single Bias. My goal is show
#the overall amount of bias in general, that is why I do not need double bias: every bias will be in a #broad category

  hate_crimes_BIAS_relevel <- hate_crimes %>% 
    separate(BIAS_DESC, into = c("BIAS_DESC"), sep = ";") %>% 
  mutate(BIAS_DESC = as_factor(BIAS_DESC)) %>% 
  mutate(BIAS_DESC =
           fct_collapse(BIAS_DESC,
                           race_ethnicity = c("Anti-Black or African American",
                                              "Anti-White",
                                              "Anti-Arab",
                                              "Anti-Asian",
                                              "Anti-Hispanic or Latino",
                                              "Anti-Other Race/Ethnicity/Ancestry",
                                              "Anti-Multiple Races, Group",
                                              "Anti-American Indian or Alaska Native",
                                              "Anti-Native Hawaiian or Other Pacific Islander"),
                        
                          sexual_orientation = c("Anti-Gay (Male)",
                                                 "Anti-Heterosexual",
                                                 "Anti-Lesbian (Female)",
                                        "Anti-Lesbian, Gay, Bisexual, or Transgender (Mixed Group)",
                                                  "Anti-Bisexual"),
                                           
                            religion = c(
                                          "Anti-Jewish",
                                          "Anti-Protestant",
                                          "Anti-Other Religion",
                                          "Anti-Islamic (Muslim)",
                                          "Anti-Catholic",
                                          "Anti-Multiple Religions, Group",
                                          "Anti-Atheism/Agnosticism",
                                          "Anti-Jehovah's Witness",
                                          "Anti-Mormon",
                                          "Anti-Buddhist",
                                          "Anti-Sikh",
                                          "Anti-Other Christian",
                                          "Anti-Hindu",
                                          "Anti-Eastern Orthodox (Russian, Greek, Other)"),
                                      
                                      disability = c("Anti-Physical Disability",
                                                      "Anti-Mental Disability"),


                                      gender_identity = c("Anti-Gender Non-Conforming",
                                                          "Anti-Transgender"),
                                      gender = c("Anti-Female",
                                                 "Anti-Male"),
                        
                               Unknown_motivation = "Unknown (offender's motivation not known)")) %>% 
  drop_na(OFFENDER_RACE)



```


	Show overview of all Hate Crimes per bias motivation of singles Bias crimes
	
```{r}
hate_crimes_BIAS_relevel %>% 
  group_by(BIAS_DESC) %>% 
  summarise(totalHatecrimes = n()) %>% 
  mutate(percentage = round(totalHatecrimes/sum(totalHatecrimes) * 100 ,2)) %>% # transform in percentage
  ggplot(aes(percentage, fct_reorder(BIAS_DESC, percentage), fill = BIAS_DESC)) + geom_col() 

```
	
	Show number of Offenders Per Race
	
```{r}
hate_crimes %>% 
  group_by(OFFENDER_RACE) %>% 
  summarise(nOfOffenders = sum(TOTAL_OFFENDER_COUNT)) %>% 
  ggplot(aes(nOfOffenders, fct_reorder(OFFENDER_RACE, nOfOffenders))) + geom_point()
 

```
 Show total numbers of crimes, by comparing 2019 and 2020, per hate crime category # I can try line graph between 2019 and 2020 with month

```{r}
hate_crimes_BIAS_relevel %>% 
  filter(year(INCIDENT_DATE) == '2019' | year(INCIDENT_DATE) == '2020') %>% 
  group_by(year = year(INCIDENT_DATE), BIAS_DESC) %>% 
  summarise(totalNumber = n()) %>% 
  ggplot(aes(BIAS_DESC, totalNumber, fill = as.factor(year) )) + geom_col(position = "dodge")


```
# Number of hate crimes each Year per Region

```{r}
 perYear_Per_region <- hate_crimes_BIAS_relevel %>%
  group_by(Year = year(INCIDENT_DATE), REGION_NAME) %>%
  summarise(totCrimes = n()) %>%
  pivot_wider(names_from = REGION_NAME,
              values_from = totCrimes)

 perYear_Per_region[c(1:5)]

```



Show Total Number of Hate Crimes per Bias where victims are groups (more than 10)
```{r}
 group_victim <- hate_crimes_BIAS_relevel %>%  filter(VICTIM_COUNT > 10) %>% count(VICTIM_COUNT,sort = TRUE)
group_victim

 group_victim_10 <- semi_join(hate_crimes_BIAS_relevel, group_victim)
 
 group_victim_10 %>% 
   group_by(BIAS_DESC) %>% 
   summarise(totNum = n()) %>% 
   ggplot(aes(fct_reorder(BIAS_DESC, totNum), totNum)) + geom_point()

```




